Exploring Alternative Routes Using Multipath TCP

### 2017-05-30 paper/presentation.tex
presentation: Implementation

\section{Implementation}
\section{Evaluation}
\section{Conclusion}

>>>>

\section{Implementation}
\SectionHead{}{
  Time to get into the actual implementation of this project.
}
\subsection{Overview}

\FigureSlide{figures/Concept.pdf}{
  \begin{itemize}
  \item Some terminology.
  \item Active opener - ``client''
  \item Passive opener - ``server''
  \item Intermediary - ``detour''
  \item This mechanism allows subflows to be created across (potentially
    multiple) detours
  \item (in addition to the main subflow directly from client to server)
  \end{itemize}
}

\begin{frame}
  \frametitle{Ingredients}

  \note{
    There are several components that go together to make this thesis work.
  }

  \begin{itemize}
  \item Multipath TCP Linux Implementation v0.91
  \item Custom path manager
  \item OpenVPN
  \item Netfilter / IPTables frameworks
  \end{itemize}
\end{frame}

\FigureSlide{figures/MovingParts.pdf}{
  \begin{itemize}
  \item There are three components that make this work.
  \item Assume server runs any MPTCP stack, unmodified
  \item First component: Client runs Linux MPTCP with custom path manager
  \item Second component: Client runs userspace daemon which communicates with
    detour.
  \item Third component: detour runs userspace daemon. Currently detour is
    restricted to any relatively modern linux (IPTables, Netfilter). Need not
    support MPTCP.
  \end{itemize}
}

\subsection{Detour Daemon}

\begin{frame}
  \frametitle{Strategies for Detours}

  \begin{itemize}
  \item OpenVPN Approach
    \begin{itemize}
    \item Establish an OpenVPN connection with detour
    \item Send packets as normal through the virtual interface
    \item Packets encapsulated via OpenVPN protocol
    \item Detour still performs NAT
    \end{itemize}
  \item NAT Approach
    \begin{itemize}
    \item Address packets directly to detour
    \item Detour alters source and destination address, forwards packet
    \item Address information must be arranged ahead of time
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{OpenVPN Approach}

  \begin{itemize}
  \item OpenVPN typically provides encryption and authentication
    \note[item]{OpenVPN is well known for providing an encrypted VPN service.
      While this is normally a good thing, in our use case, it is just
      overhead.}
  \item Configure to only provide authentication on startup, no encryption or
    message signatures
    \note[item]{To address this, we configure OpenVPN so that it will still
      require X.509 keys in order to negotiate the connection. But no encryption
      signatures will occur for data.}
  \item Use UDP as transport, to avoid ``TCP Meltdown''
    \note[item]{When TCP is encapsulated within itself, we can encounter the TCP
      Meltdown effect. This is due to multiple timers in the different layers.
      To address this, we use UDP as the transport layer.}
  \item VPN appears as network device to the kernel
  \item No per-MPTCP-connection signalling, but has per-packet overhead
    \note[item]{Some advantages here again, the VPN is just a device to the
      kernel, so this is the case MPTCP is designed to handle. There is no
      signalling, so connections can be created immediately, but there is a
      per-packet overhead.}
  \end{itemize}

  \note[item]{
    \textbf{If asked about TCP Meltdown:} Packet loss causes retransmission on
    outer TCP connection. Its retransmission timer is likely longer than the
    inner TCP. The inner timer fires a retransmission, and exponential backoff
    begins. Since the outer TCP is blocked until the retransmission, the inner
    adds up more retransmissions and exponentially backs off much more, and so
    the connection stalls quickly.
  }
\end{frame}

\begin{frame}[fragile]
  \frametitle{NAT Approach}

  \begin{center}
    \vfill
\begin{BVerbatim}
+-----------------------------------+
| ver(1) | op(1)  | reserved (2)    |
+-----------------------------------+
|           rip (4 bytes)           |
+-----------------------------------+
| rpt (2 bytes)   | dpt (2 bytes)   |
+-----------------|-----------------+
\end{BVerbatim}
    \vfill

    Custom protocol for arranging NAT detours
  \end{center}
  \note[item]{The NAT approach is perhaps the most straightforward.}
  \note[item]{The biggest issue is that a client must inform the detour of the
    server address and port. They must also agree what port on the detour to
    use. This protocol does this.}
  \note[item]{The ver field is a version, currently set to 1.}
  \note[item]{The op field is the operation, which can be either ``request'' or
    ``response''.}
  \note[item]{rip is the remote address to connect to}
  \note[item]{rpt is remote port}
  \note[item]{dpt is detour port}
  \note[item]{In the request, dpt is simply a ``proposal'', typically the same
    port as rpt. The detour replies back with the final choice of dpt.}
  \note[item]{Detour may want to choose a different one due to several types of
    port conflicts which could occur in detours, or because the proposed port
    (e.g. 80) is open and listening, and the detour doesn't want to mask it.}
\end{frame}

\subsection{Path Manager}

\begin{frame}
  \frametitle{Path Manager}

  \begin{itemize}
  \item Once a MPTCP connection is established, path manager is informed
  \item Path manager runs in a background thread
  \item Requests detours from client daemon
  \item Adds up to $N$ additional subflows, where $N$ is configurable. By
    default $N=2$
  \item Whenever a new detour becomes available, runs again
  \end{itemize}
\end{frame}

\FigureSlide{figures/KernDataStruct.pdf}{
  \begin{itemize}
  \item The path manager holds onto some internal data
  \item Tracks each NAT and VPN detour in lists, with timestamps
  \item Timestamps ensure that when the worker runs again, entries in the list
    are not reconsidered.
  \item The internal data is not global, it exists per network-namespace.
  \item Connections belong to network namespace, so do network interfaces.
  \end{itemize}
}

\subsection{Client Daemon}

\begin{frame}
  \frametitle{Client Daemon}

  \begin{itemize}
  \item Userspace daemon required for tasks which are not well-suited for the
    kernel:
    \begin{itemize}
    \item Starting processes
    \item Using UDP sockets
    \end{itemize}
  \item Daemon reads configuration file containing NAT and VPN detours.
  \item VPN instances are started up first and reported to kernel
  \item Wait for detour requests from kernel, send UDP requests, report replies
    to kernel
  \item All communication over Generic Netlink
  \end{itemize}
\end{frame}

\subsection{Putting it Together}

\begin{frame}[t]
  \frametitle{Putting it Together (NAT)}
  \begin{enumerate}
  \invisible<1>{\item[(0)] At startup, client daemon connects to VPN and reports VPN to kernel}
  \item Application creates MPTCP connection to MPTCP supporting server
  \item Once 3WHS completes, path manager requests a detour from client daemon
  \item Client daemon receives request and sends UDP request to every client
    daemon listed in configuration file
  \item Detour daemon sets up detour, sends reply
  \item Client daemon forwards reply to kernel
  \item The path manager restarts the MPTCP connection's thread, which creates
    a new subflow via this detour
  \end{enumerate}
  \note[item]{When you do it with a VPN, the main difference is that the VPN is
    already created. So the path manager can immediately create a subflow
    through the VPN, rather than waiting for a detour entry}
\end{frame}

\begin{frame}[t]
  \frametitle{Putting it Together (VPN)}

  \begin{enumerate}
  \item[(0)] At startup, client daemon connects to VPN and reports VPN to kernel
  \item Application creates MPTCP connection to MPTCP supporting server
  \item Once 3WHS completes, path manager requests a detour from client daemon.
  \item Meanwhile, it uses the VPN already available and establishes a subflow.
  \end{enumerate}
\end{frame}

\section{Evaluation}
\section{Conclusion}
